Analysis and figures

### 2025-02-17 Analysis.qmd
Update method for covariate summary

```{r Covariate_Quantity}
covariateFreq <- sheet_Covariates |>
  separate_longer_delim(Covariates, delim = ", ") |>
  separate_wider_delim(Covariates, delim = "_", 
                       names = c("Covariate", "Variation", 
                                 "Observation", "Relation"),
                       too_few = "align_start") |>
  separate_wider_delim(Covariate, delim = "-",
                       names = c("Category", "Covariate"),
                       too_few = "align_start") |>
  # Merge some covariate categories
  mutate(Covariate = case_when(Covariate == "OTHER" & Category == "E" ~ "OTHER-E",
                               Covariate == "OTHER" & Category == "S" ~ "OTHER-S",
                               Covariate %in% c("HABT", "HABA", "LACO") ~ "Habitat",
                               Covariate %in% c("CLWE", "DISA", "DIST") ~ "Climate",
                               Covariate %in% c("SPAT", "GEOM", "CONN") ~ "Spatial",
                               Covariate %in% c("ANTH", "AGRI", "PROT") ~ "Human",
                               .default = Covariate)) |>
  # Combine Review ID & Model ID, and calculate the number of studies and number of studies with each parameter
  mutate(studyID = paste(`Review ID`, `Model ID`), sep = "_") |>
  mutate(n_overall = length(unique(studyID))) |>
  select(studyID, Covariate, Category, n_overall) |>
  filter(!is.na(Covariate)) %>%
  rbind(mutate(., Covariate = Category)) %>%
  rbind(mutate(., Covariate = "ANY", Category = "ANY")) |>
  distinct() |>
  summarise(PctCategory = (n()/mean(n_overall))*100,
            .by = c(Covariate, Category)) |>
  mutate(Included = TRUE) %>%
  rbind(mutate(., PctCategory = 100-PctCategory, Included = FALSE)) |>
  mutate(Covariate = case_when(Covariate == "Habitat" ~ "Habitat",
                               Covariate == "Climate" ~ "Climate",
                               Covariate == "Spatial" ~ "Spatial",
                               Covariate == "Human" ~ "Human",
                               Covariate == "HYDR" ~ "Water",
                               Covariate == "BIOT" ~ "Biotic",
                               Covariate == "TOPO" ~ "Topography",
                               Covariate == "PHEN" ~ "Timing",
                               Covariate == "SEAS" ~ "Primary",
                               Covariate == "SURV" ~ "Secondary",
                               Covariate == "OBSE" ~ "Observation",
                               Covariate == "SITE" ~ "Site",
                               Covariate == "SPEC" ~ "Species",
                               Covariate == "ANY" ~ "Any",
                               Covariate == "E" ~ "Any\nEnv.",
                               Covariate == "S" ~ "Any\nStruct."),
         .default = Covariate) |>
  filter(!is.na(Covariate)) |>
  mutate(Covariate = fct(Covariate,
                         levels = c("Any", "Any\nEnv.",
                                    "Habitat", "Spatial", "Climate",
                                    "Timing", "Human", "Topography",
                                    "Biotic", "Water",
                                    "Any\nStruct.", "Primary", "Observation",
                                    "Secondary", "Site", "Species"))) |>
  mutate(Category = case_when(Category == "ANY" ~ "Any",
                              Category == "E" ~ "Environmental",
                              Category == "S" ~ "Structural")) |>

  ggplot(aes(x = Covariate, y = PctCategory, fill = Included)) +
  geom_col(colour = "white", 
           linewidth = 1, width = 1) +
  scale_fill_manual(limits = c(TRUE, FALSE), 
                    values = c("goldenrod", "gray50")) +
  scale_y_continuous(expand = c(0,0)) +
  facet_grid(cols = vars(Category), scales = "free_x", space = "free_x") +
  labs(y = "Percent\nof models", x = "") +
  theme(legend.position = "none",
        plot.background = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(family = "Helvetica", 
                                   angle = 90, vjust = 0.5, hjust = 1),
        strip.text.x = element_text(family = "Helvetica"))
covariateFreq
```

@@

```{r Covariate_Dynamism}
covariateDyn <- sheet_Covariates |>
  separate_longer_delim(Covariates, delim = ", ") |>
  separate_wider_delim(Covariates, delim = "_", 
                       names = c("Covariate", "Variation", 
                                 "Observation", "Relation"),
                       too_few = "align_start") |>
  separate_wider_delim(Covariate, delim = "-",
                       names = c("Category", "Covariate"),
                       too_few = "align_start") |>
  # Merge some covariate categories
  mutate(Covariate = case_when(Covariate == "OTHER" & Category == "E" ~ "OTHER-E",
                               Covariate == "OTHER" & Category == "S" ~ "OTHER-S",
                               Covariate %in% c("HABT", "HABA", "LACO") ~ "Habitat",
                               Covariate %in% c("CLWE", "DISA", "DIST") ~ "Climate",
                               Covariate %in% c("SPAT", "GEOM", "CONN") ~ "Spatial",
                               Covariate %in% c("ANTH", "AGRI", "PROT") ~ "Human",
                               .default = Covariate)) |>
  # Combine Review ID & Model ID, and calculate the number of studies and number of studies with each parameter
  mutate(studyID = paste(`Review ID`, `Model ID`), sep = "_") |>
  mutate(n_overall = length(unique(studyID))) |>
  select(studyID, `Parameter type`, Covariate, Category, Variation, n_overall) |>
  filter(!is.na(Covariate)) |>
  filter(`Parameter type` != "Initial occupancy") %>%
  rbind(rbind(mutate(., Covariate = Category),
              mutate(., Covariate = "ANY", Category = "ANY"))) |>
  mutate(n_Covariate = n(),
         .by = Covariate) |>
  mutate(Dynamism = case_when(Variation %in% c("SISE", "SU", "SE") ~ 1,
                              .default = 0)) |>
  summarise(PctDyn = (sum(Dynamism)/mean(n_Covariate))*100,
            .by = c(Covariate, Category)) |>
  mutate(Dyn = TRUE) %>%
  rbind(mutate(., PctDyn = 100-PctDyn, Dyn = FALSE)) |>
  mutate(Covariate = case_when(Covariate == "Habitat" ~ "Habitat",
                               Covariate == "Climate" ~ "Climate",
                               Covariate == "Spatial" ~ "Spatial",
                               Covariate == "Human" ~ "Human",
                               Covariate == "HYDR" ~ "Water",
                               Covariate == "BIOT" ~ "Biotic",
                               Covariate == "TOPO" ~ "Topography",
                               Covariate == "PHEN" ~ "Timing",
                               Covariate == "SEAS" ~ "Primary",
                               Covariate == "SURV" ~ "Secondary",
                               Covariate == "OBSE" ~ "Observation",
                               Covariate == "SITE" ~ "Site",
                               Covariate == "SPEC" ~ "Species",
                               Covariate == "ANY" ~ "Any",
                               Covariate == "E" ~ "Any\nEnv.",
                               Covariate == "S" ~ "Any\nStruct."),
         .default = Covariate) |>
  filter(!is.na(Covariate)) |>
  mutate(Covariate = fct(Covariate,
                         levels = c("Any", "Any\nEnv.",
                                    "Habitat", "Spatial", "Climate",
                                    "Timing", "Human", "Topography",
                                    "Biotic", "Water",
                                    "Any\nStruct.", "Primary", "Observation",
                                    "Secondary", "Site", "Species"))) |>
  mutate(Category = case_when(Category == "ANY" ~ "Any",
                              Category == "E" ~ "Environmental",
                              Category == "S" ~ "Structural")) |>

  ggplot(aes(x = Covariate, y = PctDyn, fill = Dyn)) +
  geom_col(colour = "white", 
           linewidth = 1, width = 1) +
  scale_fill_manual(limits = c(TRUE, FALSE), 
                    values = c("aquamarine3", "gray50")) +
  scale_y_continuous(expand = c(0,0)) +
  facet_grid(cols = vars(Category), scales = "free_x", space = "free_x") +
  labs(y = "Percent\ndynamic", x = "") +
  theme(legend.position = "none",
        plot.background = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(family = "Helvetica", 
                                   angle = 90, vjust = 0.5, hjust = 1),
        strip.text.x = element_blank(), strip.background = element_blank())
covariateDyn
```

@@

```{r Covariate_Observed}
covariateObs <- sheet_Covariates |>
  separate_longer_delim(Covariates, delim = ", ") |>
  separate_wider_delim(Covariates, delim = "_", 
                       names = c("Covariate", "Variation", 
                                 "Observation", "Relation"),
                       too_few = "align_start") |>
  separate_wider_delim(Covariate, delim = "-",
                       names = c("Category", "Covariate"),
                       too_few = "align_start") |>
  # Merge some covariate categories
  mutate(Covariate = case_when(Covariate == "OTHER" & Category == "E" ~ "OTHER-E",
                               Covariate == "OTHER" & Category == "S" ~ "OTHER-S",
                               Covariate %in% c("HABT", "HABA", "LACO") ~ "Habitat",
                               Covariate %in% c("CLWE", "DISA", "DIST") ~ "Climate",
                               Covariate %in% c("SPAT", "GEOM", "CONN") ~ "Spatial",
                               Covariate %in% c("ANTH", "AGRI", "PROT") ~ "Human",
                               .default = Covariate)) |>
  # Combine Review ID & Model ID, and calculate the number of studies and number of studies with each parameter
  mutate(studyID = paste(`Review ID`, `Model ID`), sep = "_") |>
  mutate(n_overall = length(unique(studyID))) |>
  select(studyID, Covariate, Category, Observation, n_overall) |>
  filter(!is.na(Covariate)) %>%
  rbind(rbind(mutate(., Covariate = Category),
              mutate(., Covariate = "ANY", Category = "ANY"))) |>
  mutate(n_Covariate = n(),
         .by = Covariate) |>
  mutate(Observed = case_when(Observation == "D" ~ 1,
                              .default = 0)) |>
  summarise(PctObs = (sum(Observed)/mean(n_Covariate))*100,
            .by = c(Covariate, Category)) |>
  mutate(Obs = TRUE) %>%
  rbind(mutate(., PctObs = 100-PctObs, Obs = FALSE)) |>
  mutate(Covariate = case_when(Covariate == "Habitat" ~ "Habitat",
                               Covariate == "Climate" ~ "Climate",
                               Covariate == "Spatial" ~ "Spatial",
                               Covariate == "Human" ~ "Human",
                               Covariate == "HYDR" ~ "Water",
                               Covariate == "BIOT" ~ "Biotic",
                               Covariate == "TOPO" ~ "Topography",
                               Covariate == "PHEN" ~ "Timing",
                               Covariate == "SEAS" ~ "Primary",
                               Covariate == "SURV" ~ "Secondary",
                               Covariate == "OBSE" ~ "Observation",
                               Covariate == "SITE" ~ "Site",
                               Covariate == "SPEC" ~ "Species",
                               Covariate == "ANY" ~ "Any",
                               Covariate == "E" ~ "Any\nEnv.",
                               Covariate == "S" ~ "Any\nStruct."),
         .default = Covariate) |>
  filter(!is.na(Covariate)) |>
  mutate(Covariate = fct(Covariate,
                         levels = c("Any", "Any\nEnv.",
                                    "Habitat", "Spatial", "Climate",
                                    "Timing", "Human", "Topography",
                                    "Biotic", "Water",
                                    "Any\nStruct.", "Primary", "Observation",
                                    "Secondary", "Site", "Species"))) |>
  mutate(Category = case_when(Category == "ANY" ~ "Any",
                              Category == "E" ~ "Environmental",
                              Category == "S" ~ "Structural")) |>

  ggplot(aes(x = Covariate, y = PctObs, fill = Obs)) +
  geom_col(colour = "white", 
           linewidth = 1, width = 1) +
  scale_fill_manual(limits = c(TRUE, FALSE), 
                    values = c("cadetblue3", "gray50")) +
  scale_y_continuous(expand = c(0,0)) +
  facet_grid(cols = vars(Category), scales = "free_x", space = "free_x") +
  labs(y = "Percent directly\nobserved", x = "") +
  theme(legend.position = "none",
        plot.background = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(family = "Helvetica", 
                                   angle = 90, vjust = 0.5, hjust = 1),
        strip.text.x = element_blank(), strip.background = element_blank())
covariateObs

@@

```{r Covariate_Complexity}
covariateComplex <- sheet_Covariates |>
  separate_longer_delim(Covariates, delim = ", ") |>
  separate_wider_delim(Covariates, delim = "_", 
                       names = c("Covariate", "Variation", 
                                 "Observation", "Relation"),
                       too_few = "align_start") |>
  separate_wider_delim(Covariate, delim = "-",
                       names = c("Category", "Covariate"),
                       too_few = "align_start") |>
  # Merge some covariate categories
  mutate(Covariate = case_when(Covariate == "OTHER" & Category == "E" ~ "OTHER-E",
                               Covariate == "OTHER" & Category == "S" ~ "OTHER-S",
                               Covariate %in% c("HABT", "HABA", "LACO") ~ "Habitat",
                               Covariate %in% c("CLWE", "DISA", "DIST") ~ "Climate",
                               Covariate %in% c("SPAT", "GEOM", "CONN") ~ "Spatial",
                               Covariate %in% c("ANTH", "AGRI", "PROT") ~ "Human",
                               .default = Covariate)) |>
  # Combine Review ID & Model ID, and calculate the number of studies and number of studies with each parameter
  mutate(studyID = paste(`Review ID`, `Model ID`), sep = "_") |>
  mutate(n_overall = length(unique(studyID))) |>
  select(studyID, Covariate, Category, Relation, n_overall) |>
  filter(!is.na(Covariate)) %>%
  rbind(rbind(mutate(., Covariate = Category),
              mutate(., Covariate = "ANY", Category = "ANY"))) |>
  mutate(n_Covariate = n(),
         .by = Covariate) |>
  mutate(Complex = case_when(Relation %in% c("N", "I", "B") ~ 1,
                               .default = 0)) |>
  summarise(PctComplex = (sum(Complex)/mean(n_Covariate))*100,
            .by = c(Covariate, Category)) |>
  mutate(Complex = TRUE) %>%
  rbind(mutate(., PctComplex = 100-PctComplex, Complex = FALSE)) |>
  mutate(Covariate = case_when(Covariate == "Habitat" ~ "Habitat",
                               Covariate == "Climate" ~ "Climate",
                               Covariate == "Spatial" ~ "Spatial",
                               Covariate == "Human" ~ "Human",
                               Covariate == "HYDR" ~ "Water",
                               Covariate == "BIOT" ~ "Biotic",
                               Covariate == "TOPO" ~ "Topography",
                               Covariate == "PHEN" ~ "Timing",
                               Covariate == "SEAS" ~ "Primary",
                               Covariate == "SURV" ~ "Secondary",
                               Covariate == "OBSE" ~ "Observation",
                               Covariate == "SITE" ~ "Site",
                               Covariate == "SPEC" ~ "Species",
                               Covariate == "ANY" ~ "Any",
                               Covariate == "E" ~ "Any\nEnv.",
                               Covariate == "S" ~ "Any\nStruct."),
         .default = Covariate) |>
  filter(!is.na(Covariate)) |>
  mutate(Covariate = fct(Covariate,
                         levels = c("Any", "Any\nEnv.",
                                    "Habitat", "Spatial", "Climate",
                                    "Timing", "Human", "Topography",
                                    "Biotic", "Water",
                                    "Any\nStruct.", "Primary", "Observation",
                                    "Secondary", "Site", "Species"))) |>
  mutate(Category = case_when(Category == "ANY" ~ "Any",
                              Category == "E" ~ "Environmental",
                              Category == "S" ~ "Structural")) |>

  ggplot(aes(x = Covariate, y = PctComplex, fill = Complex)) +
  geom_col(colour = "white", 
           linewidth = 1, width = 1) +
  scale_fill_manual(limits = c(TRUE, FALSE), 
                    values = c("plum3", "gray50")) +
  scale_y_continuous(expand = c(0,0)) +
  facet_grid(cols = vars(Category), scales = "free_x", space = "free_x") +
  labs(y = "Percent nonlinear\nor interactive", x = "") +
  theme(legend.position = "none",
        plot.background = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(family = "Helvetica", 
                                   angle = 90, vjust = 0.5, hjust = 1),
        strip.text.x = element_blank(), strip.background = element_blank())
covariateComplex
```

@@

```{r Covariate_Patchwork}
covariatePatchwork <- covariateFreq / covariateDyn / covariateObs / covariateComplex +
  plot_layout(axes = "collect") &
  theme(axis.text = element_text(family = "Helvetica"),
        axis.title = element_text(family = "Helvetica"))
covariatePatchwork

ggsave(plot = covariatePatchwork, "Figures/CovariateSummary.png")
```

>>>>

```{r Covariate_Quantity}
covariateFreq <- sheet_Covariates |>
  separate_longer_delim(Covariates, delim = ", ") |>
  separate_wider_delim(Covariates, delim = "_", 
                       names = c("Covariate", "Variation", 
                                 "Observation", "Relation"),
                       too_few = "align_start") |>
  separate_wider_delim(Covariate, delim = "-",
                       names = c("Category", "Covariate"),
                       too_few = "align_start") |>
  # Merge some covariate categories
  mutate(Covariate = case_when(Covariate == "OTHER" & Category == "E" ~ "OTHER-E",
                               Covariate == "OTHER" & Category == "S" ~ "OTHER-S",
                               Covariate %in% c("HABT", "HABA", "LACO") ~ "Habitat",
                               Covariate %in% c("CLWE", "DISA", "DIST") ~ "Climate",
                               Covariate %in% c("SPAT", "GEOM", "CONN") ~ "Spatial",
                               Covariate %in% c("ANTH", "AGRI", "PROT") ~ "Human",
                               .default = Covariate)) |>
  # Combine Review ID & Model ID, and calculate the number of studies and number of studies with each parameter
  mutate(studyID = paste(`Review ID`, `Model ID`), sep = "_") |>
  mutate(n_overall = length(unique(studyID))) |>
  select(studyID, Covariate, Category, n_overall) |>
  filter(!is.na(Covariate)) %>%
  rbind(mutate(., Covariate = Category)) %>%
  rbind(mutate(., Covariate = "ANY", Category = "ANY")) |>
  distinct() |>
  summarise(PctCategory = (n()/mean(n_overall))*100,
            .by = c(Covariate, Category)) |>
  mutate(Included = "TRUE") %>%
  rbind(mutate(., PctCategory = 100-PctCategory, Included = "FALSE")) |>
  mutate(Covariate = case_when(Covariate == "Habitat" ~ "Habitat",
                               Covariate == "Climate" ~ "Climate",
                               Covariate == "Spatial" ~ "Spatial",
                               Covariate == "Human" ~ "Human",
                               Covariate == "HYDR" ~ "Water",
                               Covariate == "BIOT" ~ "Biotic",
                               Covariate == "TOPO" ~ "Topography",
                               Covariate == "PHEN" ~ "Timing",
                               Covariate == "SEAS" ~ "Primary",
                               Covariate == "SURV" ~ "Secondary",
                               Covariate == "OBSE" ~ "Observation",
                               Covariate == "SITE" ~ "Site",
                               Covariate == "SPEC" ~ "Species",
                               Covariate == "ANY" ~ "Any",
                               Covariate == "E" ~ "Any\nEnv.",
                               Covariate == "S" ~ "Any\nStruct."),
         .default = Covariate) |>
  filter(!is.na(Covariate)) |>
  mutate(Covariate = fct(Covariate,
                         levels = rev(c("Any", "Any\nEnv.",
                                    "Habitat", "Spatial", "Climate",
                                    "Timing", "Human", "Topography",
                                    "Biotic", "Water",
                                    "Any\nStruct.", "Primary", "Observation",
                                    "Secondary", "Site", "Species")))) |>
  mutate(Category = case_when(Category == "ANY" ~ "Any",
                              Category == "E" ~ "Environmental",
                              Category == "S" ~ "Structural")) |>
  mutate(Cat_Included = paste0(Category, Included)) |>

  ggplot(aes(y = Covariate, x = PctCategory, fill = Cat_Included)) +
  geom_col(colour = "white", 
           linewidth = 1, width = 1) +
  scale_fill_manual(limits = c("AnyTRUE", "EnvironmentalTRUE", "StructuralTRUE",
                               "AnyFALSE", "EnvironmentalFALSE", "StructuralFALSE"), 
                    values = c("goldenrod", "#4D7668", "plum4",
                               "gray80", "gray80", "gray80")) +
  scale_x_continuous(expand = c(0,0), breaks = c(25,50,75)) +
  facet_grid(rows = vars(Category), 
             scales = "free_y", space = "free_y", switch = "y") +
  labs(x = "", y = "", title = "Percent of\nmodels considering:") +
  theme(legend.position = "none",
        plot.background = element_blank(),
        panel.grid = element_blank(),
        axis.text.y = element_text(family = "Helvetica"),
                                   #angle = 90, vjust = 0.5, hjust = 1),
        strip.text.y = element_text(family = "Helvetica"),
        strip.placement = "outside")
covariateFreq
```

@@

```{r Covariate_Dynamism}
covariateDyn <- sheet_Covariates |>
  separate_longer_delim(Covariates, delim = ", ") |>
  separate_wider_delim(Covariates, delim = "_", 
                       names = c("Covariate", "Variation", 
                                 "Observation", "Relation"),
                       too_few = "align_start") |>
  separate_wider_delim(Covariate, delim = "-",
                       names = c("Category", "Covariate"),
                       too_few = "align_start") |>
  # Merge some covariate categories
  mutate(Covariate = case_when(Covariate == "OTHER" & Category == "E" ~ "OTHER-E",
                               Covariate == "OTHER" & Category == "S" ~ "OTHER-S",
                               Covariate %in% c("HABT", "HABA", "LACO") ~ "Habitat",
                               Covariate %in% c("CLWE", "DISA", "DIST") ~ "Climate",
                               Covariate %in% c("SPAT", "GEOM", "CONN") ~ "Spatial",
                               Covariate %in% c("ANTH", "AGRI", "PROT") ~ "Human",
                               .default = Covariate)) |>
  # Combine Review ID & Model ID, and calculate the number of studies and number of studies with each parameter
  mutate(studyID = paste(`Review ID`, `Model ID`), sep = "_") |>
  mutate(n_overall = length(unique(studyID))) |>
  select(studyID, `Parameter type`, Covariate, Category, Variation, n_overall) |>
  filter(!is.na(Covariate)) |>
  filter(`Parameter type` != "Initial occupancy") %>%
  rbind(rbind(mutate(., Covariate = Category),
              mutate(., Covariate = "ANY", Category = "ANY"))) |>
  mutate(n_Covariate = n(),
         .by = Covariate) |>
  mutate(Dynamism = case_when(Variation %in% c("SISE", "SU", "SE") ~ 1,
                              .default = 0)) |>
  summarise(PctDyn = (sum(Dynamism)/mean(n_Covariate))*100,
            .by = c(Covariate, Category)) |>
  mutate(Dyn = "TRUE") %>%
  rbind(mutate(., PctDyn = 100-PctDyn, Dyn = "FALSE")) |>
  mutate(Covariate = case_when(Covariate == "Habitat" ~ "Habitat",
                               Covariate == "Climate" ~ "Climate",
                               Covariate == "Spatial" ~ "Spatial",
                               Covariate == "Human" ~ "Human",
                               Covariate == "HYDR" ~ "Water",
                               Covariate == "BIOT" ~ "Biotic",
                               Covariate == "TOPO" ~ "Topography",
                               Covariate == "PHEN" ~ "Timing",
                               Covariate == "SEAS" ~ "Primary",
                               Covariate == "SURV" ~ "Secondary",
                               Covariate == "OBSE" ~ "Observation",
                               Covariate == "SITE" ~ "Site",
                               Covariate == "SPEC" ~ "Species",
                               Covariate == "ANY" ~ "Any",
                               Covariate == "E" ~ "Any\nEnv.",
                               Covariate == "S" ~ "Any\nStruct."),
         .default = Covariate) |>
  filter(!is.na(Covariate)) |>
  mutate(Covariate = fct(Covariate,
                         levels = rev(c("Any", "Any\nEnv.",
                                    "Habitat", "Spatial", "Climate",
                                    "Timing", "Human", "Topography",
                                    "Biotic", "Water",
                                    "Any\nStruct.", "Primary", "Observation",
                                    "Secondary", "Site", "Species")))) |>
  mutate(Category = case_when(Category == "ANY" ~ "Any",
                              Category == "E" ~ "Environmental",
                              Category == "S" ~ "Structural")) |>
  mutate(Cat_Dyn = paste0(Category, Dyn)) |>

  ggplot(aes(y = Covariate, x = PctDyn, fill = Cat_Dyn)) +
  geom_col(colour = "white", 
           linewidth = 1, width = 1) +
  scale_fill_manual(limits = c("AnyTRUE", "EnvironmentalTRUE", "StructuralTRUE",
                               "AnyFALSE", "EnvironmentalFALSE", "StructuralFALSE"), 
                    values = c("goldenrod", "#4D7668", "plum4",
                               "gray80", "gray80", "gray80")) +
  scale_x_continuous(expand = c(0,0),  breaks = c(25,50,75)) +
  facet_grid(rows = vars(Category), scales = "free_y", 
             space = "free_y", switch = "y") +
  labs(x = "", y = "", title = "Average rate\ndynamic:") +
  theme(legend.position = "none",
        plot.background = element_blank(),
        panel.grid = element_blank(),
        axis.text.y = element_text(family = "Helvetica"), 
                                   #angle = 90, vjust = 0.5, hjust = 1),
        strip.text.y = element_blank(), 
        strip.background = element_blank(),
        strip.placement = "outside")
covariateDyn
```

@@

```{r Covariate_Observed}
covariateObs <- sheet_Covariates |>
  separate_longer_delim(Covariates, delim = ", ") |>
  separate_wider_delim(Covariates, delim = "_", 
                       names = c("Covariate", "Variation", 
                                 "Observation", "Relation"),
                       too_few = "align_start") |>
  separate_wider_delim(Covariate, delim = "-",
                       names = c("Category", "Covariate"),
                       too_few = "align_start") |>
  # Merge some covariate categories
  mutate(Covariate = case_when(Covariate == "OTHER" & Category == "E" ~ "OTHER-E",
                               Covariate == "OTHER" & Category == "S" ~ "OTHER-S",
                               Covariate %in% c("HABT", "HABA", "LACO") ~ "Habitat",
                               Covariate %in% c("CLWE", "DISA", "DIST") ~ "Climate",
                               Covariate %in% c("SPAT", "GEOM", "CONN") ~ "Spatial",
                               Covariate %in% c("ANTH", "AGRI", "PROT") ~ "Human",
                               .default = Covariate)) |>
  # Combine Review ID & Model ID, and calculate the number of studies and number of studies with each parameter
  mutate(studyID = paste(`Review ID`, `Model ID`), sep = "_") |>
  mutate(n_overall = length(unique(studyID))) |>
  select(studyID, Covariate, Category, Observation, n_overall) |>
  filter(!is.na(Covariate)) %>%
  rbind(rbind(mutate(., Covariate = Category),
              mutate(., Covariate = "ANY", Category = "ANY"))) |>
  mutate(n_Covariate = n(),
         .by = Covariate) |>
  mutate(Observed = case_when(Observation == "D" ~ 1,
                              .default = 0)) |>
  summarise(PctObs = (sum(Observed)/mean(n_Covariate))*100,
            .by = c(Covariate, Category)) |>
  mutate(Obs = "TRUE") %>%
  rbind(mutate(., PctObs = 100-PctObs, Obs = "FALSE")) |>
  mutate(Covariate = case_when(Covariate == "Habitat" ~ "Habitat",
                               Covariate == "Climate" ~ "Climate",
                               Covariate == "Spatial" ~ "Spatial",
                               Covariate == "Human" ~ "Human",
                               Covariate == "HYDR" ~ "Water",
                               Covariate == "BIOT" ~ "Biotic",
                               Covariate == "TOPO" ~ "Topography",
                               Covariate == "PHEN" ~ "Timing",
                               Covariate == "SEAS" ~ "Primary",
                               Covariate == "SURV" ~ "Secondary",
                               Covariate == "OBSE" ~ "Observation",
                               Covariate == "SITE" ~ "Site",
                               Covariate == "SPEC" ~ "Species",
                               Covariate == "ANY" ~ "Any",
                               Covariate == "E" ~ "Any\nEnv.",
                               Covariate == "S" ~ "Any\nStruct."),
         .default = Covariate) |>
  filter(!is.na(Covariate)) |>
  mutate(Covariate = fct(Covariate,
                         levels = rev(c("Any", "Any\nEnv.",
                                    "Habitat", "Spatial", "Climate",
                                    "Timing", "Human", "Topography",
                                    "Biotic", "Water",
                                    "Any\nStruct.", "Primary", "Observation",
                                    "Secondary", "Site", "Species")))) |>
  mutate(Category = case_when(Category == "ANY" ~ "Any",
                              Category == "E" ~ "Environmental",
                              Category == "S" ~ "Structural")) |>
  mutate(Cat_Obs = paste0(Category, Obs)) |>

  ggplot(aes(y = Covariate, x = PctObs, fill = Cat_Obs)) +
  geom_col(colour = "white", 
           linewidth = 1, width = 1) +
  scale_fill_manual(limits = c("AnyTRUE", "EnvironmentalTRUE", "StructuralTRUE",
                               "AnyFALSE", "EnvironmentalFALSE", "StructuralFALSE"), 
                    values = c("goldenrod", "#4D7668", "plum4",
                               "gray80", "gray80", "gray80")) +
  scale_x_continuous(expand = c(0,0),  breaks = c(25,50,75)) +
  facet_grid(rows = vars(Category), scales = "free_y", 
             space = "free_y", switch = "y") +
  labs(x = "", y = "", title = "Average rate\ndirectly observed:") + 
  theme(legend.position = "none",
        plot.background = element_blank(),
        panel.grid = element_blank(),
        axis.text.y = element_text(family = "Helvetica"), 
                                   #angle = 90, vjust = 0.5, hjust = 1),
        strip.text.y = element_blank(), 
        strip.background = element_blank(),
        strip.placement = "outside")
covariateObs

@@

```{r Covariate_Complexity}
covariateComplex <- sheet_Covariates |>
  separate_longer_delim(Covariates, delim = ", ") |>
  separate_wider_delim(Covariates, delim = "_", 
                       names = c("Covariate", "Variation", 
                                 "Observation", "Relation"),
                       too_few = "align_start") |>
  separate_wider_delim(Covariate, delim = "-",
                       names = c("Category", "Covariate"),
                       too_few = "align_start") |>
  # Merge some covariate categories
  mutate(Covariate = case_when(Covariate == "OTHER" & Category == "E" ~ "OTHER-E",
                               Covariate == "OTHER" & Category == "S" ~ "OTHER-S",
                               Covariate %in% c("HABT", "HABA", "LACO") ~ "Habitat",
                               Covariate %in% c("CLWE", "DISA", "DIST") ~ "Climate",
                               Covariate %in% c("SPAT", "GEOM", "CONN") ~ "Spatial",
                               Covariate %in% c("ANTH", "AGRI", "PROT") ~ "Human",
                               .default = Covariate)) |>
  # Combine Review ID & Model ID, and calculate the number of studies and number of studies with each parameter
  mutate(studyID = paste(`Review ID`, `Model ID`), sep = "_") |>
  mutate(n_overall = length(unique(studyID))) |>
  select(studyID, Covariate, Category, Relation, n_overall) |>
  filter(!is.na(Covariate)) %>%
  rbind(rbind(mutate(., Covariate = Category),
              mutate(., Covariate = "ANY", Category = "ANY"))) |>
  mutate(n_Covariate = n(),
         .by = Covariate) |>
  mutate(Complex = case_when(Relation %in% c("N", "I", "B") ~ 1,
                               .default = 0)) |>
  summarise(PctComplex = (sum(Complex)/mean(n_Covariate))*100,
            .by = c(Covariate, Category)) |>
  mutate(Complex = "TRUE") %>%
  rbind(mutate(., PctComplex = 100-PctComplex, Complex = "FALSE")) |>
  mutate(Covariate = case_when(Covariate == "Habitat" ~ "Habitat",
                               Covariate == "Climate" ~ "Climate",
                               Covariate == "Spatial" ~ "Spatial",
                               Covariate == "Human" ~ "Human",
                               Covariate == "HYDR" ~ "Water",
                               Covariate == "BIOT" ~ "Biotic",
                               Covariate == "TOPO" ~ "Topography",
                               Covariate == "PHEN" ~ "Timing",
                               Covariate == "SEAS" ~ "Primary",
                               Covariate == "SURV" ~ "Secondary",
                               Covariate == "OBSE" ~ "Observation",
                               Covariate == "SITE" ~ "Site",
                               Covariate == "SPEC" ~ "Species",
                               Covariate == "ANY" ~ "Any",
                               Covariate == "E" ~ "Any\nEnv.",
                               Covariate == "S" ~ "Any\nStruct."),
         .default = Covariate) |>
  filter(!is.na(Covariate)) |>
  mutate(Covariate = fct(Covariate,
                         levels = rev(c("Any", "Any\nEnv.",
                                    "Habitat", "Spatial", "Climate",
                                    "Timing", "Human", "Topography",
                                    "Biotic", "Water",
                                    "Any\nStruct.", "Primary", "Observation",
                                    "Secondary", "Site", "Species")))) |>
  mutate(Category = case_when(Category == "ANY" ~ "Any",
                              Category == "E" ~ "Environmental",
                              Category == "S" ~ "Structural")) |>
  mutate(Cat_Complex = paste0(Category, Complex)) |>

  ggplot(aes(y = Covariate, x = PctComplex, fill = Cat_Complex)) +
  geom_col(colour = "white", 
           linewidth = 1, width = 1) +
  scale_fill_manual(limits = c("AnyTRUE", "EnvironmentalTRUE", "StructuralTRUE",
                               "AnyFALSE", "EnvironmentalFALSE", "StructuralFALSE"), 
                    values = c("goldenrod", "#4D7668", "plum4",
                               "gray80", "gray80", "gray80")) +
  scale_x_continuous(expand = c(0,0), breaks = c(25,50,75)) +
  facet_grid(rows = vars(Category), scales = "free_y",
             space = "free_y", switch = "y") +
  labs(x = "", y = "", title = "Average rate non-linear\nor interactive:") +
  theme(legend.position = "none",
        plot.background = element_blank(),
        panel.grid = element_blank(),
        axis.text.y = element_text(family = "Helvetica"), 
                                   #angle = 90, vjust = 0.5, hjust = 1),
        strip.text.y = element_blank(), 
        strip.background = element_blank(),
        strip.placement = "outside")
covariateComplex
```

@@

```{r Covariate_Patchwork}
covariatePatchwork <- covariateFreq + covariateDyn + covariateObs + covariateComplex +
  plot_layout(ncol = 4, axes = "collect") &
  theme(axis.text = element_text(size = 8, family = "Helvetica", colour = "#585858",
                                 face = "italic"),
        axis.title = element_text(size = 8, family = "Helvetica", colour = "#585858"),
        plot.title = element_text(size = 8, family = "Helvetica", 
                                  colour = "#585858", hjust = 0.5)) 
covariatePatchwork

ggsave(plot = covariatePatchwork, "Figures/CovariateSummary.png",
       width = 1961, height = 1600, units = "px")
```
